Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-02-23 local-capability-machine.tex
added comment with possible crtcls implementation
adjusted the g1 proof slightly

@@

\begin{lemma}[Correctness of $g1$]
  \label{lem:correctness-g1}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{g1} & \defeq ((\entry,\glob),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{g1}) \\
    c_\stk & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    c_\link & \defeq ((\readonly,\glob),\link,\link,\link) \\
    m & \defeq \hs_{g1} \uplus 
        \ms_\flag \uplus                
        \ms_\link \uplus 
        \ms_\adv \uplus 
        \ms_\malloc \uplus 
        \ms_\stk \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{itemize}
  \item $c_\malloc$ satisfies the specification for malloc
  \end{itemize}
  \begin{align*}
    &\dom(\hs_{g1}) = [\mathtt{g1}-\olf,\mathtt{4f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{g1}(\mathtt{g1}-\olf) = ((\readonly,\glob),\link,\link,\link)$, $\ms_{g1}(\mathtt{g1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{g1}$ contains the code of $g1$ immediately followed by the code of $f4$.
  \item $\ms_\flag = [\flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\link \mapsto c_\malloc]$
  \item $\hs_\adv(\start_\adv) = c_\link$ and all other addresses of $\ms_\adv$ contain instructions.
  \item $\forall a \in \dom(\ms_\stk) \ldotp \ms_\stk(a) = 0$ %This condition could be weakened to \memSat{\ms_\stk}{[\iota^\pwl(\dom(\ms_\stk))]}
  \end{itemize}
  if 
  \[
    (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}
%
In the proof of Lemma~\ref{lem:correctness-g1}, we will use the following region
\begin{definition}
  \label{def:iotax-region}
  \begin{align*}
    \iota_x   & = (\perm, 0, \phi_\pub, \phi, H_x) \\
    \phi_\pub & = \{(0,1)\}^* \\
    \phi      & = (1,0) \union \phi_\pub \\
    H_x \; s \; \hat{W} & = \{\npair{\ms} \mid \ms(x) = s \land n > 0 \} \union \{\npair[0]{\ms}\}
  \end{align*}
\end{definition}
\begin{lemma}
  \label{lem:iotax-is-a-region}
  Definition~\ref{def:iotax-region} defines a region.
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:iotax-is-a-region}] \forcenewline
  \begin{itemize}
  \item $\phi_\pub$ is defined as the reflexive transitive closure, so it is immediately well formed.
  \item $\phi$ adds a transition to $\phi_\pub$ and is also reflexive and transitive. 
  \item $H_x$ is trivially non-expansive in the state.
  \item $H_x$ does not depend on the $\hat{W}$, so it also becomes trivially non-expansive and (privately) monotone in $\hat{W}$.
  \end{itemize}
\end{proof}

@@

          Showing
          \[
            \npair[n]{\ms_\adv} \in H^\sta_{\start_\adv,\addrend_\adv}(1)(\xi^{-1}(W_{\var{mini}}))
          \]
          is a bit more involved. It amounts to 
          \[
            \forall \addr \in \dom(\ms_\adv) \ldotp \npair[n-1]{\ms_\adv(\addr)} \in \stdvr(W_{\var{mini}})
          \]
          which in turn is trivial for everything but
          \[
            \npair[n-1]{c_\link} \in \stdvr(W_{\var{mini}})
          \]
          This amounts to showing
          \[
            \npair[n-1]{(\link,\link)} \in \readCond{}(\glob)(W_{\var{mini}})
          \]
          which amounts to
          \[
            \iota^{\sta,u} (\perma,\ms_\link) \nsubsim[n-1] \iota^\pwl_{\link,\link}
          \]
          which follows from Lemma~\ref{lem:stau-subset-pwl}.
      \end{enumproof}
      Using Lemma~\ref{lem:disj-mem-sat} repeatedly with \ref{pf:g1-3-1}, \ref{pf:g1-3-2}, and \ref{pf:g1-3-3} gives the desired memory satisfaction.
    \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$ \label{pf:g1-4}\\
      This amounts to showing
      \begin{enumproof}
        \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}(c_\stk)} \in \stdvr(W)$ \\
          The assumptions on $c_\stk$ and $\ms_\stk$ in the lemma entail
          \begin{itemize}
          \item \pointstostack{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}}{\emptyset}{\ms_\stk}
          \end{itemize}
          and further there is a $\iota^\pwl$ region for $\ms_\stk$ in $W$, so the result follows from Lemma~\ref{lem:stack-cap-vr}.
        \item $\npair{c_{g1}} \in \stdvr(W)$ \\
          Let $n_1 < n$ and $W_1 \futurestr W$ and show
          \[
            \npair[n_1]{\updatePcPerm{c_{g1}}} \in \stder(W_1)
          \]
          To this end assume $n_2 \leq n_1$, $\memSat[n_2]{\ms_1}{W_1}$, and $\npair[n_2]{\reg_1} \in \stdrr(W_1)$ and show
          \[
            \npair[n_2]{(\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1)} \in \observations(W_1)
          \]
          % TODO elaborate on what happens here using anti reduction lemma and malloc spec.
          Use Lemma~\ref{lem:anti-red-obs}, so let $\ms_f'$ be given, then
          \[
            (\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1 \uplus \ms_f') \step[i] (\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls \uplus \ms_f')
          \]
          for some $i$. Where
          \begin{itemize}
          \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc'}{W_1\update{0}{\iota_\malloc}} $
          \item $\iota_\malloc \futurewk \iota_{\malloc,0}$
          \item $\ms_x = [x \mapsto 0]$
          \item $c_x = ((\rwx,\glob),x,x,x)$
          \item $\ms_\env = [\env \mapsto c_x]$
          \item $c_\env = ((\rwx,\glob,\env,\env,\env))$
          \item $c_{f4}$ is $c_{g1}$ updated to point at $f4$.
          \item $\ms_\cls$ contains $c_\env$, the $c_{f4}$, and the activation code.
          \item $c_\cls$ is a global enter capability for $\ms_\cls$ pointing to the activation code.
          \item $\reg_2 = reg_0\update{\pcreg}{\updatePcPerm{\reg_1(r_0)}}\update{r_0}{\reg_1(r_0)}\update{r_1}{c_\cls}$
          \end{itemize}
          For world
          \[
            W_2 = W_1\update{0}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms_\env \uplus \ms_\cls)}\update{i_2}{\iota_x}
          \]
          where $i_1,i_2\not\in\dom(W_1)$ and $i_1 \neq i_2$ and $\iota_x$ is the region in Definition~\ref{def:iotax-region} which is a region by Lemma~\ref{lem:iotax-is-a-region}.
          Show
          \begin{equation}
            \label{eq:g1-obs-to-show}
            \npair[n_2-i]{(\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls)} \in \observations(W_2)
          \end{equation}
          If $\reg_1(r_0).\perm \not\in \{\entry, \exec, \rwx, \rwlx\}$, then the execution fails after the jump and \label{eq:g1-obs-to-show} is thus trivially true.

@@

            \item $\npair[n_2-i]{\reg_2} \in \stdrr(W_2)$ \\
              Amounts to showing
              \begin{enumproof}
                \item $\npair[n_2-i]{\reg_2(r_0)} \in \stdvr(W_2)$ by assumption $\npair[n_2]{\reg_1} \in \stdrr(W_1)$ and $\stdvr$ monotonicity wrt. $\futurewk$
                \item $\npair[n_2-i]{c_\cls} \in \stdvr(W_2)$ \\
                  Let $n_3 < n_2 - i$ and $W_3 \futurestr W_2$ be given and show
                  \[
                    \npair[n_3]{\updatePcPerm{c_\cls}} \in \stder(W_3)
                  \]
                  To this and let $n_4 \leq n_3$, $\memSat[n_4]{\ms_3}{W_3}$, and $\npair[n_4]{\reg_3} \in \stdrr(W_3)$ and show
                  \begin{equation}
                    \label{eq:g1-c-cls-obs}
                    \npair[n_4]{(reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3)} \in \observations(W_3)
                  \end{equation}
                  Let $\ms_3^p$ and $\ms_3^t$ be memory segments such that $\ms_3 = \ms_3^p \uplus \ms_3^t$ and $\memSat[n_4]{\ms_3^p}{\revokeTemp{W_3}}$ (using Lemma~\ref{lem:priv-mono-like2})

                  Use Lemma~\ref{lem:anti-red-obs}, let $\ms_f''$ be given and take $\ms_r$ to be the part of $\ms_3^t$ that $\reg_3(r_\stk)$ does not govern. By the operational semantics, we know\footnote{the execution may fail, but then the configuration is trivially in the observation relation.}
                    (reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3^p \uplus \ms_3^t \uplus \ms_f'') \step[j] (\reg_4,\ms_4 \uplus \ms_3^t \uplus \ms_f'')

@@

                    \item \texttt{prepstack} did not fail, so the capability must be $\rwlx$ and follow the stack convention.
                  \item $\reg_3(r_1)$ is a $\glob$ capability.

@@

                  \item $\reg_3(r_\env) = c_\env$

@@

                    \memSat[n_4-j]{\ms_4}{W_4}

@@

                    \npair[n_4-j]{(\reg_4,\ms_4 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused)} \in \observations(W_4)

@@

                    \item $\memSat[n_4-j]{\ms_4}{\revokeTemp{W_4}}$: follows from $W_4 = \revokeTemp{W_4}$

@@

                        \npair[n_4-j]{\updatePcPerm{reg_3(r_1)}} \in \stder(W_5)

@@

                        \item $\memSat[n_4-j-1]{\ms_5}{W_5}$ for some $\ms_5$ which is one of the assumptions of Hyp-Callee.
                        \item $\npair[n_4-j-1]{\reg_5} \in \stdrr(W_5)$ where $\reg_5$ is as described in the \texttt{scall} lemma Hyp-callee premise.\\

@@

                          1) $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_5)$, use Lemma~\ref{lem:stdvr-glob-priv-mono} with $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_3)$, the capability is global, and $W_5 \futurestr W_3$.

@@

                        \npair[n_4-j-1]{(\reg_5,\ms_5)} \in \observations(W_5)

@@

                      \item $n_5 \leq n_4-j-2$

@@

\subsubsection{Create closure macro-instruction correctness}
\begin{lemma}[$\mathtt{crtcls}$ works]
  If
  \begin{itemize}
  \item \lookingat{(\reg,\ms)}{\mathtt{crtcls}\;\overline{(x,r)}\;r}{c_{\var{next}}}
  \item \linksto{(\reg,\ms)}{\malloc}{k}{c_\malloc}
  \item $c_\malloc$ satisfies the $\malloc$ specification with $\iota_{\malloc,0}$
  \item $W \futurestr [i \mapsto\iota_{\malloc,0}]$
  \item $\memSat{\ms}{W}$
  \item $\ms = \ms' \uplus \ms_{\var{footprint}}$
  \item $\memSat{\ms_{\var{footprint}}}{[i \mapsto W(i)]}$
  \item[Hyp-Cont]
    If
    \begin{itemize}
    \item $n' \leq n$
    \item $\iota_\malloc \futurewk W(i)$
    \item $\memSat[n']{\ms' \uplus \ms_{\var{footprint}}'}{W[i \mapsto \iota_\malloc]}$
    \item $\memSat{\ms_{\var{footprint}}'}{[i \mapsto \iota_\malloc]}$
    \item
      \begin{equation*}
        \reg'(r') =
        \begin{cases}
          c_{\var{next}} & r' = \pcreg \\
          c_\cls = ((\entry,\glob),\start,\addrend,\start+2) & r' = r_1 \\
          \reg(r)      & r' \not\in \{\pcreg, r_1\} \union \RegName_t
        \end{cases}
      \end{equation*}
    \item $\ms_\cls = \ms_\act \uplus \ms_\env$
    \item $c_\env = ((\readwrite,\glob),\start_\env,\addrend_\env,\start_\env)$
    \item $\dom(\ms_\env) = [\start_\env,\addrend_\env]$
    \item $\ms_\env(\start_\env,\dots,\addrend_\env) = \reg(\overline{r})$
    \item Hyp-act\\ If
      \begin{itemize}
      \item $\reg''(\pcreg) = \updatePcPerm{c_\cls}$
      \end{itemize}
      Then  $\exists k \ldotp \forall \ms_f \ldotp (\reg'',\ms'' \uplus \ms_\cls \uplus \ms_f) \step[k] (\reg''',\ms'' \uplus \ms_\cls \uplus \ms_f)$
      where
        \begin{equation*}
          \reg'''(r') =
          \begin{cases}
            c_\env & r' = c_\env \\
            \updatePcPerm{\reg(r)} & r' = \pcreg \\
            \reg''(r') & r' \not\in \RegName_t
          \end{cases}
        \end{equation*}
    \end{itemize}
    Then we have $\npair[n']{(\reg',\ms' \uplus \ms_{\var{footprint}} \uplus \ms_\cls)} \in \observations(W[i \mapsto \iota_\malloc])$
  \end{itemize}
  Then 
  \[
    \npair{(\reg,\ms)} \in \observations(W)
  \]
\end{lemma}

>>>>

%Possible implementation of crtcls \overline{(x,r_v)} r_code where |\overline{(x,r_v)}| = n (i_1,...,i_6 under)
% malloc r_t1 n
% store r_t1 r_v1
% lea r_t1 1
% store r_t1 r_v2
% lea r_t1 1
% ...
% lea r_t1 1
% store r_t1 r_vn
% lea r_t1 -n
% restrict r_t1 encodePerm(rw)
% malloc r_1 8 //length of activation record
% store r_1 r_code
% lea r_1 1
% store r_1 r_t1
% move r_t1 0
% lea r_1 1
% store r_1 encode(i_1)
% lea r_1 1
% store r_1 encode(i_2)
% lea r_1 1
% ...
% lea r_1 1
% store r_1 encode(i_6)
% lea r_1 -5 //offset to first instruction
% restrict r_1 encodePerm(e)

%Activation code (i_1,...,i_6)
% i_1 = move r_t1 pc
% i_2 = lea r_t1 -2
% i_3 = load r_env r_t1
% i_4 = lea r_t1 1
% i_5 = load r_t1 r_t1
% i_6 = jmp r_t1

@@

\begin{lemma}[Correctness of $g1$]
  \label{lem:correctness-g1}
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
    c_{g1} & \defeq ((\entry,\glob),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{g1}) \\
    c_\stk & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
    c_\link & \defeq ((\readonly,\glob),\link,\link,\link) \\
    m & \defeq \hs_{g1} \uplus 
        \ms_\flag \uplus                
        \ms_\link \uplus 
        \ms_\adv \uplus 
        \ms_\malloc \uplus 
        \ms_\stk \uplus
        \ms_{\var{frame}} 
  \end{align*}
  where 
  \begin{itemize}
  \item $c_\malloc$ satisfies the specification for malloc with $\iota_{\malloc,0}$
  \end{itemize}
  \begin{align*}
    &\dom(\hs_{g1}) = [\mathtt{g1}-\olf,\mathtt{4f}] \\
    &\dom(\hs_\flag) = [\flag,\flag] \\
    &\dom(\ms_\link) = [\link,\link]\\
    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
  \end{align*}
  and
  \begin{itemize}
  \item $\ms_{g1}(\mathtt{g1}-\olf) = ((\readonly,\glob),\link,\link,\link)$, $\ms_{g1}(\mathtt{g1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{g1}$ contains the code of $g1$ immediately followed by the code of $f4$.
  \item $\ms_\flag = [\flag \mapsto 0]$
  \item $\ms_{\var{link}} = [\link \mapsto c_\malloc]$
  \item $\hs_\adv(\start_\adv) = c_\link$ and all other addresses of $\ms_\adv$ contain instructions.
  \item $\forall a \in \dom(\ms_\stk) \ldotp \ms_\stk(a) = 0$ %This condition could be weakened to \memSat{\ms_\stk}{[\iota^\pwl(\dom(\ms_\stk))]}
  \end{itemize}
  if 
  \[
    (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step[n] (\halted,m'),
  \]
  then
  \[
    m'(\flag) = 0
  \]  
\end{lemma}
%
In the proof of Lemma~\ref{lem:correctness-g1}, we will use the following region
\begin{definition}
  \label{def:iotax-region}
  \begin{align*}
    \iota_x   & = (\perm, 0, \phi_\pub, \phi, H_x) \\
    \phi_\pub & = \{(0,1)\}^* \\
    \phi      & = (1,0) \union \phi_\pub \\
    H_x \; s \; \hat{W} & = \{\npair{\ms} \mid \ms(x) = s \land n > 0 \} \union \{\npair[0]{\ms}\}
  \end{align*}
\end{definition}
\begin{lemma}
  \label{lem:iotax-is-a-region}
  Definition~\ref{def:iotax-region} defines a region.
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:iotax-is-a-region}] \forcenewline
  \begin{itemize}
  \item $\phi_\pub$ is defined as the reflexive transitive closure, so it is immediately well formed.
  \item $\phi$ adds a transition to $\phi_\pub$ and is also reflexive and transitive. 
  \item $H_x$ is trivially non-expansive in the state.
  \item $H_x$ does not depend on the $\hat{W}$, so it also becomes trivially non-expansive and (privately) monotone in $\hat{W}$.
  \end{itemize}
\end{proof}

@@

          Showing
            \npair[n]{\ms_\adv} \in H^\sta_{\start_\adv,\addrend_\adv}(1)(\xi^{-1}(W_{\var{mini}}))
          is a bit more involved. It amounts to 
            \forall \addr \in \dom(\ms_\adv) \ldotp \npair[n-1]{\ms_\adv(\addr)} \in \stdvr(W_{\var{mini}})
          which in turn is trivial for everything but
            \npair[n-1]{c_\link} \in \stdvr(W_{\var{mini}})
            \npair[n-1]{(\link,\link)} \in \readCond{}(\glob)(W_{\var{mini}})
          which amounts to
            \iota^{\sta,u} (\perma,\ms_\link) \nsubsim[n-1] \iota^\pwl_{\link,\link}
          which follows from Lemma~\ref{lem:stau-subset-pwl}.
      Using Lemma~\ref{lem:disj-mem-sat} repeatedly with \ref{pf:g1-3-1}, \ref{pf:g1-3-2}, and \ref{pf:g1-3-3} gives the desired memory satisfaction.
    \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$ \label{pf:g1-4}\\
        \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}(c_\stk)} \in \stdvr(W)$ \\
          The assumptions on $c_\stk$ and $\ms_\stk$ in the lemma entail
          \item \pointstostack{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}}{\emptyset}{\ms_\stk}
          and further there is a $\iota^\pwl$ region for $\ms_\stk$ in $W$, so the result follows from Lemma~\ref{lem:stack-cap-vr}.
        \item $\npair{c_{g1}} \in \stdvr(W)$ \\
          Let $n_1 < n$ and $W_1 \futurestr W$ and show
            \npair[n_1]{\updatePcPerm{c_{g1}}} \in \stder(W_1)
          To this end assume $n_2 \leq n_1$, $\memSat[n_2]{\ms_1}{W_1}$, and $\npair[n_2]{\reg_1} \in \stdrr(W_1)$ and show
            \npair[n_2]{(\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1)} \in \observations(W_1)
          Using Lemma~\ref{lem:malloc-works}, Lemma~\ref{lem:crtcls-works}, Lemma~\ref{lem:anti-red-obs} (and some others),
          it suffices to show
            \npair[n_2']{(\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\cls \uplus \ms_x)} \in \observations(W_2)
            W_2 = W_1\update{0}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms_\cls)}\update{i_2}{\iota_x}
          where $i_1,i_2\not\in\dom(W_1)$ and $i_1 \neq i_2$ and $\iota_x$ is the region in Definition~\ref{def:iotax-region} which is a region by Lemma~\ref{lem:iotax-is-a-region}. Also
          \item $\iota_\malloc \futurestr \iota_\malloc'$
          \item $c_x = ((\rwx,\glob),x,x,x)$
          \item $\ms_x = [x \mapsto 0]$
          \item $\memSat[n_2']{\ms_2 \uplus \ms_\malloc' \uplus \ms_\cls \uplus \ms_x}{W_2}$
          \item $c_\env = ((\rwx,\glob,\env,\env,\env))$
          \item $\ms_\env = [\env \mapsto c_x]$
          \item $c_{f4} = ((\rwx,\glob),\mathtt{g1} - \olf,\mathtt{4f},\mathtt{f4})$
          \item $\ms_\cls = \ms_\env \uplus \ms_\act$
              \reg_2(r) =
                \updatePcPerm{\reg_1(r_0)} & r = \pcreg \\
                \reg_1(r_0) & r = r_0 \\
                c_\cls & r = r_1 \\
                0 & \text{otherwise}
          Finally assume Hyp-Act:
            \begin{multline}
            \label{eq:g1-hyp-act}
            \forall \reg,\ms \ldotp \reg(\pcreg)=c_\cls \Rightarrow\\
            \exists j \ldotp \forall \ms_f \ldotp (\reg,\ms \uplus \ms_\cls \uplus \ms_f) \step[j] (\reg\update{\pcreg}{\updatePcPerm{c_{f4}}}\update{r_\env}{c_\env},\ms \uplus \ms_\cls \uplus \ms_f)
          \end{multline}
            \label{eq:g1-obs-to-show}
            \npair[n_2-i]{(\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls)} \in \observations(W_2)
          If $\reg_1(r_0).\perm \not\in \{\entry, \exec, \rwx, \rwlx\}$, then the execution fails after the jump and \label{eq:g1-obs-to-show} is thus trivially true.

@@

            \item $\npair[n_2-i]{\reg_2} \in \stdrr(W_2)$ \\
              Amounts to showing
                \item $\npair[n_2-i]{\reg_2(r_0)} \in \stdvr(W_2)$ by assumption $\npair[n_2]{\reg_1} \in \stdrr(W_1)$ and $\stdvr$ monotonicity wrt. $\futurewk$
                \item $\npair[n_2-i]{c_\cls} \in \stdvr(W_2)$ \\
                  Let $n_3 < n_2 - i$ and $W_3 \futurestr W_2$ be given and show
                    \npair[n_3]{\updatePcPerm{c_\cls}} \in \stder(W_3)
                  To this and let $n_4 \leq n_3$, $\memSat[n_4]{\ms_3}{W_3}$, and $\npair[n_4]{\reg_3} \in \stdrr(W_3)$ and show
                    \label{eq:g1-c-cls-obs}
                    \npair[n_4]{(reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3)} \in \observations(W_3)
                  Let $\ms_3^p$ and $\ms_3^t$ be memory segments such that $\ms_3 = \ms_3^p \uplus \ms_3^t$ and $\memSat[n_4]{\ms_3^p}{\revokeTemp{W_3}}$ (using Lemma~\ref{lem:priv-mono-like2}). By $\memSat[n_4]{\ms_3}{W_3}$ and $W_3 \futurestr W_2$, we know $\ms_\cls \subseteq \ms_3^p$, so using Hyp-Act(\ref{eq:g1-hyp-act}), we get $j$ such that
                    \forall \ms_f \ldotp (reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3^p \uplus \ms_3^t \uplus \ms_f) \step[j] (\reg_3\update{\pcreg}{\updatePcPerm{c_\cls}}\update{r_\env}{c_\env},\ms_3^p \uplus \ms_3^t \uplus \ms_f)
                  Using Lemma~\ref{lem:anti-red-obs} it suffices to show
                    \npair[n_4]{(\reg_3\update{\pcreg}{\updatePcPerm{c_\cls}}\update{r_\env}{c_\env},\ms_3^p \uplus \ms_3^t)} \in \observations(W_3)
                  Use Lemma~\ref{lem:anti-red-obs} again. This time let $\ms_f''$ be given and take $\ms_r$ to be the part of $\ms_3^t$ that $\reg_3(r_\stk)$ does not govern. By the operational semantics, we know\footnote{the execution may fail, but then the configuration is trivially in the observation relation.}
                    (\reg_3\update{\pcreg}{\updatePcPerm{c_\cls}}\update{r_\env}{c_\env},\ms_3^p \uplus \ms_3^t \uplus \ms_f'') \step[j']
                    (\reg_4,\ms_4 \uplus \ms_3^t \uplus \ms_f'')

@@

                    \item \texttt{prepstack} did not fail, so the stack capability must be $\rwlx$ and follow the stack convention.
                  \item $\reg_4(r_1)$ is a $\glob$ capability.

@@

                  \item $\reg_4(r_\env) = c_\env$

@@

                    \memSat[n_4-j-j']{\ms_4}{W_4}

@@

                    \npair[n_4-j-j']{(\reg_4,\ms_4 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused)} \in \observations(W_4)

@@

                    \item $\memSat[n_4-j-j']{\ms_4}{\revokeTemp{W_4}}$: follows from $W_4 = \revokeTemp{W_4}$

@@

                        \npair[n_4-j-j']{\updatePcPerm{reg_3(r_1)}} \in \stder(W_5)

@@

                        \item $\memSat[n_4-j-j'-1]{\ms_5}{W_5}$ for some $\ms_5$ which is one of the assumptions of Hyp-Callee.
                        \item $\npair[n_4-j-j'-1]{\reg_5} \in \stdrr(W_5)$ where $\reg_5$ is as described in the \texttt{scall} lemma Hyp-callee premise.\\

@@

                          1) $\npair[n_4-j-j'-1]{\reg_3(r_1)} \in \stdvr(W_5)$, use Lemma~\ref{lem:stdvr-glob-priv-mono} with $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_3)$, the capability is global, and $W_5 \futurestr W_3$.

@@

                        \npair[n_4-j-j'-1]{(\reg_5,\ms_5)} \in \observations(W_5)

@@

                      \item $n_5 \leq n_4-j-j'-2$

@@

\subsubsection{Create closure macro-instruction correctness}
\begin{lemma}[$\mathtt{crtcls}$ works]
  \label{lem:crtcls-works}
  If
  \begin{itemize}
  \item \lookingat{(\reg,\ms)}{\mathtt{crtcls}\;\overline{(x,r)}\;r}{c_{\var{next}}}
  \item \linksto{(\reg,\ms)}{\malloc}{k}{c_\malloc}
  \item $c_\malloc$ satisfies the $\malloc$ specification with $\iota_{\malloc,0}$
  \item $W \futurestr [i \mapsto\iota_{\malloc,0}]$
  \item $\memSat{\ms}{W}$
  \item $\ms = \ms' \uplus \ms_{\var{footprint}}$
  \item $\memSat{\ms_{\var{footprint}}}{[i \mapsto W(i)]}$
  \item[Hyp-Cont]
    If
    \begin{itemize}
    \item $n' \leq n$
    \item $\iota_\malloc \futurewk W(i)$
    \item $\memSat[n']{\ms' \uplus \ms_{\var{footprint}}'}{W[i \mapsto \iota_\malloc]}$
    \item $\memSat{\ms_{\var{footprint}}'}{[i \mapsto \iota_\malloc]}$
    \item
      \begin{equation*}
        \reg'(r') =
        \begin{cases}
          c_{\var{next}} & r' = \pcreg \\
          c_\cls = ((\entry,\glob),\start,\addrend,\start+2) & r' = r_1 \\
          \reg(r)      & r' \not\in \{\pcreg, r_1\} \union \RegName_t
        \end{cases}
      \end{equation*}
    \item $\ms_\cls = \ms_\act \uplus \ms_\env$
    \item $c_\cls = ((\entry,\glob),\dots)$
    \item $c_\env = ((\readwrite,\glob),\start_\env,\addrend_\env,\start_\env)$
    \item $\dom(\ms_\env) = [\start_\env,\addrend_\env]$
    \item $\ms_\env(\start_\env,\dots,\addrend_\env) = \reg(\overline{r})$
    \item Hyp-act\\ If
      \begin{itemize}
      \item $\reg''(\pcreg) = \updatePcPerm{c_\cls}$
      \end{itemize}
      Then  $\exists k \ldotp \forall \ms_f \ldotp (\reg'',\ms'' \uplus \ms_\cls \uplus \ms_f) \step[k] (\reg''',\ms'' \uplus \ms_\cls \uplus \ms_f)$
      where
        \begin{equation*}
          \reg'''(r') =
          \begin{cases}
            c_\env & r' = c_\env \\
            \updatePcPerm{\reg(r)} & r' = \pcreg \\
            \reg''(r') & r' \not\in \RegName_t
          \end{cases}
        \end{equation*}
    \end{itemize}
    Then we have $\npair[n']{(\reg',\ms' \uplus \ms_{\var{footprint}} \uplus \ms_\cls)} \in \observations(W[i \mapsto \iota_\malloc])$
  \end{itemize}
  Then 
  \[
    \npair{(\reg,\ms)} \in \observations(W)
  \]
\end{lemma}
